---
title: "Assignment Umair Amjad 46462252"
output:
  pdf_document: 
    latex_engine : xelatex
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
 
#q1a)
```{r}
mr_dat = read.table("surg.dat", header = TRUE)
head(mr_dat)
plot(mr_dat[,-6], pch=19)





```

the gender variable needs to be removed as gender is a categorical variable and it can alter the results of the scatterplot significantly
the plot sugests that the dots are significantly close to each suggesting that the data isnt evenly spread.


By analysisng the matrix we can conclude that survival has a stronger relationship with enyzme and liver. Survival has moderate relationship with blood and week relationship with age

# q1b)
```{r}
mr_dat = read.table("surg.dat", header = TRUE)
round(cor(mr_dat[,-6]),2)

```
Therefore we can see liver has a better relationship with the survival variable while on the other hand,  the worst relationship is between the variables age and enyzme 



# q1c)

``` {r}
mr_dat = read.table("surg.dat", header = TRUE)
mylm = lm(survival~. , data = mr_dat)
summary(mylm)
anova(mylm)





```
 
```{r}
mylm8= lm(survival~ blood+prognosis+ enzyme, data= mr_dat)
anova(mylm8)

```
 survival = -11.367 + 86.630(blood)+8.501(prognosis)+11.125(Enzyme)+38.507(Liver)-2.340(age)-0.2201(gender)

 $H_0: \beta_1 = \beta_2.......\beta_k\$ 
 $H_1: \beta_i\ne0$
 

therefore due to the pvalue being 1.19e-10 therefore we have evidence to  reject this null hypothesis as 1.19e-10<0.05 and the f statistic is 17.85 with the degree ofs freedom between 6 and 47. The anova table suggests that the pvalue of blood,prognesis,enyzme is less then 0.05 while on the other hand the pvalues of  liver,age and gender is greater then 0.05 therefore these variables shouldnt be included in the equation.
for the anova table liver,age,gender pvalue >0.05 meaning that these variable needs to be remove from the anova table and once the variable were remove from the anova table blood pvalue increase, enyzme pvalue increase while the enyzme pvalue got smaller.
therefore is a relationship between aall the predictors


# q1d)


```{r}
mylm2 = lm(survival~ blood+prognosis+enzyme+liver+age, data= mr_dat)
summary(mylm2)



```

```{r}
mylm3 = lm(survival~ blood+prognosis+enzyme+age, data= mr_dat) 
summary(mylm3)



```

final model
```{r}

mylm4= lm(survival~ blood+prognosis+enzyme, data = mr_dat)
summary(mylm4)          



```


this model demonstrated above is the best model as the variable gender isnt included due to how gender had a big influences on the model
as the summary lm show that liver,age gender needs to be remove from the model as it exceeds the 0.05 significant level so therefore the equation should include survival, enzyme,blood,prognosis
survival = -1410.847 +101.054 (blood)+ 9.382(prognosis)+ 12.128(Enzyme)

#q1 e 
```{r}
par( mfrow = c(1,2))
mylm6<-lm(survival~ blood+prognosis+enzyme, data=mr_dat)
plot(mylm6$fitted, mylm6$residuals, main = "Residuals vs Fitted",
xlab = "Fitted", ylab = "Residuals")
qqnorm(mylm6$residuals, main = "Normal Q-Q plot of residuals")


```

the normal q-q plots seems  to have a close linear action demonstrating that there are errors as the datas gets close to normally distributed. The residuals vs fitted doesnt have a pattern.So therefore the regression model isnt approate model to this case due to these reasons

# q1 f
```{r}

mylm5 = lm(log(survival)~ blood+prognosis+enzyme+liver+age , data = mr_dat)
summary(mylm5)



``` 

```{r}
mylm7= lm(log(survival)~ blood+prognosis+enzyme+age, data = mr_dat)
summary(mylm7)

```

```{r}
mylm8= lm(log(survival)~ blood+prognosis+enzyme, data = mr_dat)
summary(mylm8)


```




#q1 g 

If the regression model isnt approiate in terms of survival then the next best option is to utilized log therefore log is the best option for survival response. The log method removes all the outliers and make relations more linear

#Q2 a

```{r}
car_data = read.table("kml.dat", header=TRUE)
 table(car_data[,c("car","driver")])


``` 
This data seems to be balanced due to how it shows the equals number of observations for each variable 

#Q2b
```{r}
 par(mfrow =  c(1,2))
 with(car_data,interaction.plot(driver,car,kmL,trace.label = "driver", xlab= "car", ylab = "kmL"))
 with(car_data,interaction.plot(car,driver,kmL,trace.label="car",xlab = "driver", ylab= "kmL"))

```

```{r}

boxplot(kmL ~ driver + car, data = car_data)



```


Also the prelimary investigation suggest that the lines are parrall meaning that therefore there is no interaction between the lines meaning that the boxplot can be easily intrepret. Therefore the interecation seems to be insignificant.


#Q2c
```{r}
summary(aov(kmL~  driver* car, data=car_data))





```

$H_0: \beta_1 = \beta_2.......\beta_k\ 
$H_1: \beta_i\ne0$



Through the anova table results we can conclude that the pvalues results is less then 0.05 therefore these results are insignifcant. The assumption in this case can be concluded that there is no interectation between the data as considered before the lines dont intersect, and the pvalues is less then 0.05




#2d
The preliminary investigation suggests that the lines are parrall meaning that the interecation is insignifcant this can conclude that the anova results conclude that the pvalues for car and driver is less then 0.05.Therefore those variables shouldnt be removed for this case.



